The promise of community-driven preprints in ecology and evolution

### 2024-10-03 ms/ms.qmd
Some text edits

*EcoEvoRxiv* is one of the few community-driven preprint servers that has paved the way for new initiatives, by accepting multilingual preprints, registered reports, and non-traditional research reports. Such initiatives are distinct from other preprint servers, such as bioRxiv, which only accepts empirical research in English. *EcoEvoRxiv* also promotes peer review and community discussion in the hopes of improving the quality of preprints and speeding up their peer-reviewed publication (e.g., Peer Community In -- PCI [@guillemaud2019peer]). Peer Community In allows for fast, constructive peer review around a preprint with peer reviews being transparent and published online [@guillemaud2019peer].

@@

  # Subset the preprints
  preprints <- data2  %>% filter(time_between_preprint_and_pub_days >= 60)

  dates_preprints <- range(preprints$preprint_published_date)
##---------------------##
# Submissions to  EcoEvoRxiv over time (everything)
##---------------------##
  angle = 45
  p1 <- ggplot(data2, aes(x = preprint_pub_date)) +
    geom_histogram(fill = "white", color = "black") +
    scale_y_continuous(limits = c(0, 130), expand = c(0, 0)) +
    labs(x = "Published date", y = "Number of papers") +
    theme_classic() +
    #theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    theme(plot.tag = element_text(size = 16, face = "bold")) + 
      # Soft Launch of EcoEvoRxiv
    annotate( "segment", x = as.POSIXct("2018-06-01"),  xend = as.POSIXct("2018-06-01"), y = 0, yend = 82, color = "black", linetype = "dashed", size = 0.5) + 
    annotate("text", x = as.POSIXct("2018-06-01"), y = 85, label = "Soft Launch", size = 4, angle = angle, hjust = 0) +
    # Launch of EcoEvoRxiv
    annotate( "segment", x = as.POSIXct("2018-11-11"),  xend = as.POSIXct("2018-11-11"), y = 0, yend = 82, color = "black", linetype = "dashed", size = 0.5) + 
    annotate("text", x = as.POSIXct("2018-11-11"), y = 85, label = "Launch", size = 4, angle = angle, hjust = 0) +
    # Spanish and Port
    annotate( "segment", x = as.POSIXct("2023-02-14"),  xend = as.POSIXct("2023-02-14"), y = 0, yend = 82, color = "black", linetype = "dashed", size = 0.5) + 
    annotate("text", x = as.POSIXct("2022-11-14"), y = 85, label = "Spanish and Portug.", size = 4, angle = angle, hjust = 0) +
    # Transition to CDL
    annotate( "segment", x = as.POSIXct("2022-09-14"),  xend = as.POSIXct("2022-09-14"), y = 0, yend = 82, color = "black", linetype = "dashed", size = 0.5) + 
    annotate("text", x = as.POSIXct("2022-07-14"), y = 85, label = "CDL", size = 4, angle = angle, hjust = 0) +
   # IUCN Red Eco submission 29 August 2023
   # 1000 preprints
    annotate( "segment", x = as.POSIXct("2023-08-29"),  xend = as.POSIXct("2023-08-29"), y = 0, yend = 82, color = "black", linetype = "dashed", size = 0.5) + 
    annotate("text", x = as.POSIXct("2023-04-29"), y = 85, label = "IUCN Report", size = 4, angle = angle, hjust = 0)

@@

Research can take a while to be published (see below). However, Early and Mid-Career Researchers (EMCRs) (~10 years post-PhD) are under pressure to publish rapidly to be competitive in job applications, promotions, and obtaining grants to progress their careers [@vale2015accelerating; @sarabipour2019value]. Preprints are one way EMCRs can achieve faster dissemination and greater visibility [@fu2019releasing]. As such, EMCRs may be expected to make use of preprints more than colleagues at later career stages. We collected data on the 'academic age' of submitting authors by looking at Google Scholar profiles of authors (when available) and recording their first year of publication in a peer-reviewed journal. While this is a rough estimate of career stage, there was evidence that the number of preprints posted decreases with later career stages (negative binomial glm: year slope = `r res[2, "Estimate"]`, SE: `r round(res[2, "Std. Error"], 2)`, *p* < 0.001, n = `r nrow(count_dat[complete.cases(count_dat),])` years). Most preprints were submitted by authors who published their first paper in the last ~10 years ([@fig-summary]E), with the median year since first publication being `r median(data2$submitting_author_first_publication_year, na.rm = TRUE)` (mean = `r mean(data2$submitting_author_first_publication_year, na.rm = TRUE)`; SD = `r sd(data2$submitting_author_first_publication_year, na.rm = TRUE)`, n = `r sum(complete.cases(data2$submitting_author_first_publication_year))`). These patterns support the expectation that EMCRs may use preprints to make their work more visible and disseminate their findings more quickly. However, we acknowledge that validating this conclusion does require more rigorous experimental approaches.

>>>>

*EcoEvoRxiv* is one of the few community-driven preprint servers that has paved the way for new initiatives, by accepting multilingual preprints, registered reports, and non-traditional research reports. Such initiatives are distinct from other preprint servers, such as bioRxiv, which only accept empirical research in English. *EcoEvoRxiv* also promotes peer review and community discussion in the hopes of improving the quality of preprints and speeding up their peer-reviewed publication (e.g., Peer Community In -- PCI [@guillemaud2019peer]). Peer Community In allows for fast, constructive peer review around a preprint with peer reviews being transparent and published online [@guillemaud2019peer].

@@

# Dataframes
            submissions <- data2          # Cntains all submissions, preprints, postprints and all article types
              preprints <- data2  %>% filter(time_between_preprint_and_pub_days >= 60) # Contains only preprints and not postprints
preprints_research_only <- data2  %>% filter(time_between_preprint_and_pub_days >= 60 & type_of_preprint == "research article") # Limits preprints to research only articles

  # Subset the preprints
  dates_preprints <- range(preprints$preprint_published_date)
##---------------------##
# Submissions to  EcoEvoRxiv over time (everything)
##---------------------##
  angle = 45
  p1 <- ggplot(data2, aes(x = preprint_pub_date)) +
    geom_histogram(fill = "white", color = "black") +
    scale_y_continuous(limits = c(0, 130), expand = c(0, 0)) +
    labs(x = "Published date", y = "Number of papers") +
    theme_classic() +
    #theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    theme(plot.tag = element_text(size = 16, face = "bold")) + 
      # Soft Launch of EcoEvoRxiv
    annotate( "segment", x = as.POSIXct("2018-06-01"),  xend = as.POSIXct("2018-06-01"), y = 0, yend = 82, color = "black", linetype = "dashed", size = 0.5) + 
    annotate("text", x = as.POSIXct("2018-06-01"), y = 85, label = "Soft Launch", size = 4, angle = angle, hjust = 0) +
    # Launch of EcoEvoRxiv
    annotate( "segment", x = as.POSIXct("2018-11-11"),  xend = as.POSIXct("2018-11-11"), y = 0, yend = 82, color = "black", linetype = "dashed", size = 0.5) + 
    annotate("text", x = as.POSIXct("2018-11-11"), y = 85, label = "Launch", size = 4, angle = angle, hjust = 0) +
    # Spanish and Port
    annotate( "segment", x = as.POSIXct("2023-02-14"),  xend = as.POSIXct("2023-02-14"), y = 0, yend = 82, color = "black", linetype = "dashed", size = 0.5) + 
    annotate("text", x = as.POSIXct("2022-11-14"), y = 85, label = "Spanish and Portug.", size = 4, angle = angle, hjust = 0) +
    # Transition to CDL
    annotate( "segment", x = as.POSIXct("2022-09-14"),  xend = as.POSIXct("2022-09-14"), y = 0, yend = 82, color = "black", linetype = "dashed", size = 0.5) + 
    annotate("text", x = as.POSIXct("2022-07-14"), y = 85, label = "CDL", size = 4, angle = angle, hjust = 0) +
   # IUCN Red Eco submission 29 August 2023
   # 1000 preprints
    annotate( "segment", x = as.POSIXct("2023-08-29"),  xend = as.POSIXct("2023-08-29"), y = 0, yend = 82, color = "black", linetype = "dashed", size = 0.5) + 
    annotate("text", x = as.POSIXct("2023-04-29"), y = 85, label = "IUCN Report", size = 4, angle = angle, hjust = 0)

@@

Research can take a while to be published (see below). Early and Mid-Career Researchers (EMCRs) (~10 years post-PhD) are under pressure to publish rapidly to be competitive in job applications, promotions, and obtaining grants to progress their careers [@vale2015accelerating; @sarabipour2019value]. Preprints are especially useful for EMCRs because they can achieve faster dissemination and greater visibility [@fu2019releasing]. As such, EMCRs may be expected to make use of preprints more than colleagues at later career stages. We collected data on the 'academic age' of submitting authors by looking at Google Scholar profiles of authors (when available) and recording their first year of publication in a peer-reviewed journal. While this is a rough estimate of career stage, there was evidence that the number of preprints posted decreases with later career stages (negative binomial glm: year slope = `r res[2, "Estimate"]`, SE: `r round(res[2, "Std. Error"], 2)`, *p* < 0.001, n = `r nrow(count_dat[complete.cases(count_dat),])` years). Most preprints were submitted by authors who published their first paper in the last ~10 years ([@fig-summary]E), with the median year since first publication being `r median(data2$submitting_author_first_publication_year, na.rm = TRUE)` (mean = `r mean(data2$submitting_author_first_publication_year, na.rm = TRUE)`; SD = `r sd(data2$submitting_author_first_publication_year, na.rm = TRUE)`, n = `r sum(complete.cases(data2$submitting_author_first_publication_year))`). These patterns support the expectation that EMCRs may use preprints to make their work more visible and disseminate their findings more quickly. However, we acknowledge that validating this conclusion does require more rigorous experimental approaches.
